---
title: "Lab 6"
author: "Katheryn Moya"
date: "11/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)
```

## Rank based test example

Mann-Whitney U (non-parametric alternative to an unpaired 2-sample t-test).

H0: Equal ranks (or medians)
HA: Rank are not equal (or, medians are not equal)

### Create samples

`set.seed()` to create "pseudorandom, but reproducible samples"


```{r}
# set a starting point so that random sample can be reproducible

set.seed(1414)
gp_1 <- sample.int(20, size = 15, replace = TRUE)

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE)
```

Take a quick look at the distributions using `hist()`

```{r}
hist(gp_1)
hist(gp_2)
```

Try a t-test so we can compare the results:

```{r}

ttest <- t.test(gp_1, gp_2)

# If these data were paired, I can just add the argument `paired - "TRUE"` within the t-test
```

This p-value means that IF our samples are drawn from populations with the same mean (i.e., if the null hypothesis is true), there is `r round(ttest$p.value,2)` probably of finding sample means *at least as different* as mine by random chance. --- Pretty likely that these populations came from samples with the same mean -> so you would accept null hypothesis, because you don't think it's that unlikely that it could happen randomly -- fail to reject null hypothesis

[A more comprehensice explanation of differences between groups... actually means, data spread, effect size]...there is no significant difference (t(df) = t-statistic value, p = ).

### Mann Whitney U rank based test

We use the `wilcox.text()` function to perform a two-sample rank-based test.

```{r}
mwu <- wilcox.test(gp_1,gp_2)

# for this rank based test, 
```

This p-value means that if our samples are drawn from a population with the same median, there is a 28% chance of finding sample medians at least as different as ours by random chance. 

No significant difference is medians (Mann-Whitney U, U = `r mwu$statistic`, p = `r round(mwu$p.value, 2)`)

Example text for reporting: "The medians of group A (M = 22.4 g) and group B (M = 31.2 g) did not differ significantly (Mann Whitney U, ...)"

## Simple linear regression

Two *continuous quantitative* variables that you are tryin gto understand the relationship between, and you think that a linear relationship makes sense.

Use the `penguins` data from {palmerpenguins}, exploring flipper length (x) versus body mass (y)

```{r}
# Use View(penguins) to look at it 
# make an exploratory scatterplot of penguin flipper length (x-axis) versus body mass (y-axis)

ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point() +
  geom_smooth(method = "lm")
```

### Find the model estimates (coefficients)

Use the `lm()` function for linear model.

```{r}
penguins_lm <- lm(body_mass_g ~ flipper_length_mm, data = penguins)

```
 
 ^^^tells me that body mass in grams of penguins is modeled by 49.69 times flipper length, minus 5780.8
 
Model coefficients: `r penguins_lm$coefficients`
Flipper length coefficient: `r penguin_lm$coefficients[2]`

### Use `broom::tidy()` to get a tidy output from lm()

```{r}

penguins_lm_tidy <- tidy(penguins_lm)

penguins_lm_tidy

```
### Use broom::glance() to get output of model fit and statistical significance

```{r}
penguins_lm_glance <- glance(penguins_lm)

penguins_lm_glance

```

### Use equatiomatic::extract_eq() to return model in LaTeX

```{r}

extract_eq(model = penguins_lm, use_coefs = TRUE)
```

 ### Use plot() function to automatically produce diagnostic plots
 
```{r}
plot(penguins_lm)
```
 
 
 
 

